---
title: "p8105_hw2_bh2852"
author: "Beicheng Huang"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r load_libraries}
library(tidyverse)
library(readxl)
```
### Problem 1

We clean the 538 `pols` data, which provides information on the number
of national politicians who are democratic or republican at any given
time. There are some values for which `prez_gop` is `2` – these are
months in which Ford became President following Nixon’s resignation. In
the new `president` variable created as part of our data cleaning, we
code these as `gop` (same as values when `prez_gop` is `1`).

```{r}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```

We also clean the 538 `snp` data, which contains information related to
Standard & Poor’s stock market index.

``` {r}
snp = 
  read_csv(
    "./data/fivethirtyeight_datasets/snp.csv",
    col_types = cols(date = col_date(format = "%m/%d/%y"))) |>
  separate(date, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    year = if_else(year > 2023, year - 100, year)) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, close) 
```

Finally, we tidy the `unemployment` data so that it can be merged with
the `pols` and `snp` datasets.

``` {r}
unemployment = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

Merge the three datasets

``` {r}
data_538 = 
  left_join(pols, snp) |>
  left_join(x = _, y = unemployment)

str(data_538)
## tibble [822 × 13] (S3: tbl_df/tbl/data.frame)
##  $ year        : num [1:822] 1947 1947 1947 1947 1947 ...
##  $ month       : chr [1:822] "January" "February" "March" "April" ...
##  $ month_num   : int [1:822] 1 2 3 4 5 6 7 8 9 10 ...
##  $ gov_gop     : num [1:822] 23 23 23 23 23 23 23 23 23 23 ...
##  $ sen_gop     : num [1:822] 51 51 51 51 51 51 51 51 51 51 ...
##  $ rep_gop     : num [1:822] 253 253 253 253 253 253 253 253 253 253 ...
##  $ gov_dem     : num [1:822] 23 23 23 23 23 23 23 23 23 23 ...
##  $ sen_dem     : num [1:822] 45 45 45 45 45 45 45 45 45 45 ...
##  $ rep_dem     : num [1:822] 198 198 198 198 198 198 198 198 198 198 ...
##  $ president   : chr [1:822] "dem" "dem" "dem" "dem" ...
##  $ month_abb   : chr [1:822] "Jan" "Feb" "Mar" "Apr" ...
##  $ close       : num [1:822] NA NA NA NA NA NA NA NA NA NA ...
##  $ unemployment: num [1:822] NA NA NA NA NA NA NA NA NA NA ...
```


## Problem 2

Import and Clean Mr Trash Wheel
```{r import mr}
# import data from excel with specify sheet number = "Mr. Trash Wheel", omit column with notes
mr_trash_wheel = read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = cell_cols("A2:N586")) %>%
#clean name
  janitor::clean_names() %>% 
# omit rows do not include dumpster-specific data
  drop_na(dumpster) %>% 
# create new home powered variable
  mutate(homes_powered = weight_tons*500/30,name = "Mr.TrashWheel",year = as.numeric(year)) %>% 
  relocate(name,dumpster,year,month) 

mr_trash_wheel
```

```{r import prof}
professor_trash_wheel = read_xlsx( 
  "202309 Trash Wheel Collection Data.xlsx",
  cell_cols("A2:M108"),
  sheet = "Professor Trash Wheel") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(homes_powered = weight_tons*500/30,
         name = "Prof.TrashWheel") %>% 
  relocate(name,dumpster,year,month) 

professor_trash_wheel
```

```{r import gwynnda}
gwynnda_trash_wheel = read_xlsx( 
  "202309 Trash Wheel Collection Data.xlsx",
  cell_cols("A2:L157"),
  sheet = "Gwynnda Trash Wheel") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(homes_powered = weight_tons*500/30,
         name = "Gwynnda.TrashWheel") %>% 
  relocate(name,dumpster,year,month) 

gwynnda_trash_wheel
```


```{r Combine}
Combined_dataset <- bind_rows(mr_trash_wheel, professor_trash_wheel, gwynnda_trash_wheel)
nrow(Combined_dataset)
ncol(Combined_dataset)

Combined_dataset
```

There are 845 observations and 15 variables in the Combined dataset.

```{r P.total weightton}
filter(Combined_dataset, name == "Prof.TrashWheel") %>% 
  pull(weight_tons) %>% 
  sum()
```
The total weighttons for professor trash wheel is 216.26

```{r Gwynnda.Cigaratte_butts}
filter(Combined_dataset, name == "Gwynnda.TrashWheel", year == 2021, month == "July")%>% 
  pull(cigarette_butts)%>% 
  sum()
```
The total number of cigarette butts collected by Gwynnda in July of 2021 is 16300


Key Variables:
name:identifier for the Trash Wheel (Mr.TrashWheel, Prof.Trash Wheel, or Gwynnda.TrashWheel).
date: The date of data collection.
homes_powered: The number of homes powered, if applicable.
total_weight: The total weight of trash collected.
cigarette_butts: The count of cigarette butts collected.



## Problem 3

Import,clean the baseline data
```{r}
baseline_data = read.csv(file = "MCI_baseline.csv", skip=1, na = c(".")) %>%
  janitor::clean_names() %>%
  mutate(sex = factor(sex, labels = c("Female", "Male")),        
         apoe4 = factor(apoe4, labels = c("APOE4 non-carrier", "APOE4 carrier")),        
         age_at_onset = as.numeric(age_at_onset))

baseline_data
```
Exclude participant with MCI before enrollment
```{r}
filtered_data = baseline_data %>%
  filter(age_at_onset - current_age > 0 | is.na(age_at_onset)) 

filtered_data
```
Exclude participant who did not develop MCI
```{r}
participated_data = filtered_data %>%
  filter(age_at_onset - current_age > 0) %>% 
  drop_na(age_at_onset)
    
participated_data
```


Calculate mean of age in final participated data
```{r}
round(mean(filtered_data$current_age, na.rm=TRUE),digits = 0)

```
Calculate the percentage of women who are APOE4 carrier
```{r}
nrow(filter(filtered_data, sex == "Female" & apoe4 == "APOE4 carrier")) / nrow(filter(filtered_data, sex == "Female"))
```

Before cleaning, there are 479 participants.In the filtered dataset(exclude participant who developed MCI before enrollment), there are 93 participants who developed MCI(those non-qualified were excluded) in the final participated dataset. 
By using round(mean(filtered_data$current_age, na.rm=TRUE),digits = 0), we have the average MCI onset age as 65. 
By using nrow(filter(filtered_data, sex == "Female" & apoe4 == "APOE4 carrier")) / nrow(filter(filtered_data, sex == "Female")), we get there are 30% of female participants having MCI.


##Amyloid data load 

```{r amyloid data loading}
amyloid_data = read.csv(file = "mci_amyloid.csv", skip=1) %>%
  janitor::clean_names() %>%
  pivot_longer(baseline:time_8,
               names_prefix="time_",
               names_to="years_since_diagnosed",
               values_to="apoe4_42/40 ratio") %>%
  mutate(years_since_diagnosed=replace(years_since_diagnosed, years_since_diagnosed=="baseline", 0)) %>%
  mutate(years_since_diagnosed=as.numeric(years_since_diagnosed))

amyloid_data
```
The amyloid dataset contains 5 observations for all 483 participants which results in 2435 observations.


```{r Finding unique ID}
baseline <- filtered_data %>%
  anti_join(amyloid_data, by = c("id" = "study_id")) %>%
  select(id)

amyloid <- amyloid_data %>%
  anti_join(filtered_data, by = c("study_id" = "id")) %>%
  select(study_id)
```
There are 8 observations in baseline dataset with unique ID.
There are total of 80 observations in amyloid which consist of 16 unique id with 5 observations each.

```{r Combining datasets where ID are presented in both sets}
Combined_MCI_amy = 
  inner_join(filtered_data,amyloid_data, by = c("id" = "study_id"))
```
After combining by ID presented in both sets, there are a total of 2355 observations which consists of 471 participants(ID presented in both sets) with 5 data entries each.

##Export
```{r}
# Export the result as a CSV
write_csv(Combined_MCI_amy, "./Result.csv")
```